Analytics Without Parameter Tuning Considered Harmful?

### 2015-03-13 learners.tex
Updates from ShareLaTeX

Dta miners find generalizations (a.k.a. summaries)
of the data. All such generalizations must be smaller than the data it generalizes (otherwise we might as well just keep the original data and ignore the generalization)\footnote{Technical note:
even instance-based reasoners uses generalizations when they partially match new cases to old;
or if they use some  clustering algorithm to restrict the case matching
to just the subset of old cases most similar  to new  cases.}. 
Generalization is important since,
without generalization, all we can do is perform exact matches
of new situations to a database of old examples. This is not recommended practice since,
 if the new situation has not occurred  before, then nothing in the data  will
 matches the new situation
(and we cannot make any predictions).

The key point here is that there are many ways to summarize data and exceConsider some learner that
builds predictive rules form a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into five ranges, then the 20 attributes of \fig{ck} can be combined into 
$2^{5*20}> 10^{30}$
possible rules

>>>>

This section describes the learners used in this study. While there are many differences
in those learners, they have the same goal: to summarize a large data set into a small model\footnote{Technical note: proponents of instance-based reasoning might argue that
their goal is to retain, and not summarize, the training data.
However, 
even instance-based reasoners use some degree of summarizing when they learn
ways to bias the instance matching away from irrelevant training data.
Those bias tools include  generalizing points to ranges (to allow for partial matches)
or  clustering algorithm that restrict the case matching
to just the subset of old cases most similar  to new  cases.}.

For example,
consider some learner that
builds predictive rules from a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into $d=5$ divisions, then the 20 attributes of \fig{ck} can be combined into 
$2^{A*d} = 2^{5*20} > 10^{30}$
possible rules. The key thing to note here is that many of these rules will have similar effects
since, at least in software engineering, data sets containing rows $R$ and columns $C$ can contain much redundancy:
\bi
\item When $c \subset C$ columns of data that are correlated to each other, then those
$c$ columns can be replaced by one column.
\item When $r \subset R$ rows of data are very similar (e.g. they come from similar projects), 
then those $r$ rows can be replaced
by one cluster centroid that stands in $r$.
\item
Kocagenunil~\cite{me13a} and Papakroni~\cite{papa13} conduct  different experiments
to  show  that tables of software defect data in the form of \fig{ck} contain around 90\% redundant
data. Specifically, they show that such tables can be be pruned back to around
25\% of the columns and 10\% of the rows without losing the essential information in the data set.
\ei
Given this high level of redundancy, learners need only explore a small fraction of these $2^{A*d}$ rules.
Most learners exploit this fact to optimize their reasoning. Rather than explore all options,
they employ some {\em search bias} that lets them quickly focus on a small set of promising rules.

Different biases work best for different data sets so it is standard practice to allow an analyst
to control these biases via tuning parameters. For example, consider the search biases
$B_i$ of decision tree learner that :
\bi
\item Splits the data on some criteria $B_1$;
\item Assesses each split using some criteria $B_2$;
\item Selects which split(s) to explore further using $B_3$;
\item Recurses into each split until some stopping criteria $B_4$ is encountered.
\ei
These biases also appear when the learned decision tree makes a decision:
\bi
\item To declare a software module to be XXX
\ei

### 2015-03-13 tuningpredictor.tex
Updates from ShareLaTeX

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{ 
Analytics Without Parameter Tuning Considered Harmful?}
\numberofauthors{1}
\author{
\alignauthor
Wei Fu, Tim Menzies, Vivek Nair\\
       \affaddr{Computer Science, North Carolina State University, Raleigh, USA}\\
       fuwei.ee, tim.menzies, vivekaxl@gmail.com}

\maketitle
\begin{abstract}
One of the ``black arts'' of data mining is setting the tuning
parameters that control   choices within a data miner.  We offer a simple,
automatic, and very effective  method for finding those tunings.

@@

One of the ``black arts'' of data mining is setting the tuning
parameters that control  the choices within that data miner.
Many researchers have applied automatic optimizers to find good tunings.
For example, 
the search-based SE community has   developed techniques
for tuning that treating parameters  as a configuration
search space~\cite{cora10,Krogmann10}. More generally,
the field of {\em hyper-heuristics} explores methods for auto-adapting
options within the device searching for solutions~\cite{jia2013learning}.

To the best of our knowledge, this paper is the first exploration in software engineering
of applying automatic optimizers to the task of tuning data miners for defect prediction
(though see~\cite{cora10,balogh12} for work on effort estiamtion and genetic optimizers).
We shows that for at least one software analytics problem
(software defect prediction from static code attributes),   a
very simple optimizer called 
{\em differential evolution} can  find good tunings.

The good news here is that  tuning defect predictors is very easy and very effective. 
Differential
evolution is so simple to implement that many
researchers just ``roll their own'' rather
than relying on some standard package.
By various measures
(presented below),  
tuning defect predictors is at least an order of magnitude
simpler than other standard optimization tasks. 
Also, measured in terms on runtimes, 
differental evolution takes  seconds
to a just few minutes to propose highy effective tunings.
Further, the tunings found by differential evolution
can have a dramatic change to the performance of a learner;
e.g. in this paper we show that tuning can alter the precision of
a software defect predictor from 2\% to 98\%.

The not-so-good-news is that these results  challenge much prior work into defect prediction.
Several   prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task~.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
softwre defect predition~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests). Hence, all those prior results that ranked learners for software
analytics now need to be revisited and perhaps revised.

Another class of conclusions challenged by this paper are research papers
that use data miners to   show that certain factors
are more inflential than others for (say)
predicting defects. As shown below, such conclusions can be dramatically
changed by the tuning process since those  ``influential'' factors are very different pre- and post- tuning. Also, those factors tend to  change from project to project or if the goal
of the tuning is altered.
Hence, many old papers    need to be revisted  and perhaps revised~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting}.  
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted since that
Menzies study did conduct a tuning study before reporting what factors the data miners
found where most influential.

In summary,it is now an open and pressing research issue to check if
analytics without parameter tuning is considered {\em harmful} or, at the 
very least, {\em misleading}.
At least for defect prediction,
it is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.
There are many other kinds of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc) and the implication of this
study for those tasks is unclear. 
However,  those other tasks often use the same kinds of learners
explored in this paper so it is clearly possible that
the conclusions of this paper apply to other SE analytics tasks as well.

@@

Two caveats before beginning. Firstly, we are {\em not} saying that {\em all learning requires
tuning}. For example, when proposing fixes to software, Weimar et al. use a genetic learner
to propose patches~\cite{Weimer:2009}. Suppose  the performance criteria for that work is how many known bugs can be found
and fixed automatically. Note that there is no {\em better than} in that performance criteria so
that study does not need tuning. However, once {\em better than} enters the performance criteria
then this becomes a race between competing methods or attributes.  In such a race, it is unfair
to hobble one competitor with poor tunings. Hence, we recommend tuning for  all the data mining
studies mentioned above\cite{lessmann2008benchmarking,hall11,me07b,bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting} since
they are arguing that one learner or one set of attributes
is {\em better than} another.

Secondly,
One caveat before beginning- this paper explores
optimization via differential evolution.
XXX

>>>>

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{ Analytics Without Parameter Tuning Considered Harmful?}
\numberofauthors{1}
\author{\alignauthor Wei Fu, Tim Menzies, Vivek Nair\\
       \affaddr{Computer Science, North Carolina State University, Raleigh, USA}\\
       fuwei.ee, tim.menzies, vivekaxl@gmail.com}
\maketitle
\begin{abstract}
One of the ``black arts'' of data mining is setting the tuning
parameters that control   choices within a data miner.  We offer a simple,
automatic, and very effective  method for finding those tunings.

@@

One of the ``black arts'' of data mining is setting the tuning
parameters that control  the choices within that data miner.
Many researchers have applied automatic optimizers to find good tunings.
The search-based SE community has   developed techniques
for tuning that treat parameters  as a configuration
search space~\cite{cora10,Krogmann10}. 
The field of {\em hyper-heuristics} explores methods for auto-adapting
options within the device searching for solutions~\cite{jia2013learning} in applications
like test-case generation and code refactorings.

To the best of our knowledge, this paper is the first extensive exploration 
of applying automatic optimizers to tune data miners for defect prediction,
(though see~\cite{cora10,balogh12,Minku13,minku13z} for tuning effort estimators).
The one similar study  we can find in the period 2004 to 2014\footnote{See the GECCO and SSBSE proceedings at    goo.gl/2MY602 and goo.gl/uzvU8e.} was
Bouktiff et al.~\cite{Bouktif06} who used  simulated annealing to tune Bayes classifiers.
That was a very limited study, applied to a  single defect data set.
The analysis of this paper is far more extensive and offers the following novel result:
\bi
\item
Tuning static code defect predictors is {\em remarkably simple} and can {\em dramatically improve the performance}
of those learners. 
\ei
Prior to these experiments, our expectation was tuning
would be an  extensive and expensive evolutionary optimization procedure. 
A standard run of such evolutionary optimizers requires   thousands,
if not millions, of evaluations.
To
our surprise, we achieved dramatic improvements in the performance scores
of our data miners after  mere 50 to 80 evaluations (!!) of a very simple evolutionary 
optimizer  called differential evolution~\cite{storn1997differential}.
Better yet,  those  tunings (found so quickly)   
  have a dramatic change to the performance of a learner. For example,
in this paper we show that tuning can alter the precision of
a software defect predictor from 2\% to 98\%.

Our tuning results  challenge much prior work in software analytics. 
Firstly, there exist research papers
that use data miners to   show that certain factors
are more influential than others for (say)
predicting defects. As shown below, such conclusions can be dramatically
changed by the tuning process since those  ``influential'' factors are very different pre- and post- tuning. Also, those factors tend to  change from project to project or if the goal
of the tuning is altered.
Hence, many old papers    need to be revisted  and perhaps revised~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting}.  
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted since that
Menzies study did conduct a tuning study before reporting what factors the data miners
found where most influential.

Secondly, several   prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task~.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
software defect prediction~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests). Hence, all those prior results that ranked learners for software
analytics now need to be revisited and perhaps revised.

Thirdly, it is standard practice to use the default ``off-the-shelf'' tunings  for data mining tools (previously
we have defended that approach on methodological grounds arguing that it
encourages reproducibility~\cite{me15:book1}). That ``off-the-shelf''  policy
can no longer be condoned. For example, one such default setting
  in the Python SciKitLearn toolkit~\cite{scikit-learn}
is to use $T=10$ decision trees in  Random Forest classifiers.
Our optimizer settled on  $T$ values that were were nowhere near that default:

{\scriptsize
\[T \in \left\{\begin{array}{l} 55,  65, 70,   82, 88, 96, 100,  102,  104, 107,\\
                                108,  119, 133,  140, 140,   147,  145,  142   \end{array}\right\}
\]}

In summary,it is now an open and pressing research issue to check if
analytics without parameter tuning is considered {\em harmful} or, at the 
very least, {\em misleading}.
At the very least, there is now cause to doubt many conclusions based on
``off-the-shelf'' tunings.
At the very most, 
it is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.

@@

\section{Preliminaries}

Before beginning, this section offers certain caveats on these results.

Firstly, we are {\em not} saying that {\em all} learning requires
tuning. For example, when proposing fixes to software, Weimar et al. use a genetic learner
to propose patches~\cite{Weimer:2009}. Consider one performance criteria for that work; i.e. 
that   that that method can find and fix known bugs. Note that this is a {\em competency criteria}
which does not include the phrase  {\em better than}. Such performance criteria do
not requiring tuning.  However, once {\em better than} enters the performance criteria
(as done in \cite{lessmann2008benchmarking,hall11,me07b,bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting})
then this becomes a race between competing methods (or attributes).  In such a race, it is unfair
to hobble one competitor with poor tunings.

Secondly, the tuning results shown here only came from one particular software analytics task 
(defect prediction from static code attributes).
There are many other kinds of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc) and the implication of this
study for those tasks is unclear. 
However,  those other tasks often use the same kinds of learners
explored in this paper so it is quite possible that
the conclusions of this paper apply to other SE analytics tasks as well.

Thirdly, this paper explores {\em some} learners using {\em one}  optimizer. Hence, it makes
no claim that this is the {\em best} optimizer for {\em all} learners.
Rather, our point is that there exists at least some learners
whose performance can be dramatically improved by 
at least one simple optimization scheme.  We hope that this work inspires
much future work as this community develops and debugs best practices for tuning
software analytics.

Fourthly, it would be incorrect to say that this paper is arguing that
software analytics is somehow wrong-headed, misguided, and we should not do it anymore.
In the age of the Internet and global access to software engineering data,
there exists the  problem of information overload. {\em Something} must be done to
allow analysts to make conclusions via an automatic analysis over a lot of data.
The results of this paper is that for a particular local context
(a specific data set and a specific goal) there exists  
methods for optimizing the conclusions reached in that context.  Those conclusions
may not generalize to other contexts but this  is not a council for despair. While there may
not exist general conclusions, there does seem to exist general methods for finding
local conclusions in a particular context. Further, as shown below, those
methods may be very simple to implement and very fast to execute.
